---
title: "Alpine Phenology"
author: "M. DePrenger-Levin and RA Hufft"
date: "August 28, 2015"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
---

If you need these packages
`install.packages("RCurl")`      
`install.packages("ggplot2")`     
`install.packages("data.table")`
'install.packages("curl")'
install.packages("Rcpp")

#Try a `rm(list=ls())` before running

```{r}
library(RCurl)
library(ggplot2)
library(data.table)
library(raster)
require(rgdal)
library(devtools)
# devtools::install_github("ropensci/prism")
library(prism)
```

Bring csv in from the website
I get an error: `Error in str.default(obj, ...) : invalid multibyte string 2` but will see if it matters...   
looks like it's the lat long that didn't get in correctly
```{r, echo=FALSE}
bloom<-read.csv(text = getURL("https://raw.githubusercontent.com/DenverBotanicGardens/alpine-phenology/master/QR_final_R_plant_climate_data_2.csv"))

head(bloom)
```

If not working to load from github:
```{r}
#bloom<-read.csv(path.expand("P:/alpine-phenology/QR_final_R_plant_climate_data.csv"))

#Bring in Q/All.Projects_by_species/aa_Spapefiles_Maps/aa_QGIS Projects/AlpinePhenologyproject/

#bloom<-read.csv(path.expand("P:/alpine-phenology/TB_tempPrecipData.csv"),        
#                header = TRUE, as.is=TRUE)   
```

# Add growing degree days   
White etal 2015     
Cumulative Growing Degree Days 
$\sum_{i=1}^{n} \left(T_{mean} - T_{base})$   
$T_{base}$ should be the lowest air temperature when flowering will occur    
Temperature Threshold and GDD models...
    
    
In the csv Tbase is 10, not negative 10. Which is correct?     
 McMaster and Wilhelm 1997:      
Tbase = -10    
Corn $T_{base} = 10$   
Winter wheat $T_{base} = 0$   
temperature below which the process of interest does not progress
$$GDD = \frac{T_{MAX} + T_{MIN}}{2} - T_{BASE}$$   


# not really GDD... 
```{r}
Tbase <- -10
bloom$GDD2 <- (bloom$Avg_Hi + bloom$Avg_Lo)/2 - Tbase 

#cumsum(bloom$GDD2) # need by day not by species, same data per species


```


```{r}
#created new column that divides Raw_precip by 100 (see PRISM documentation) to get mm
head(bloom)
names(bloom)
str(bloom)  #tells me column name, data type and examples, errors in lat long when brought in from web

#Climate data figures
#This section runs linear regression on Avg High temp by Year
#hist(bloom$Avg_Hi)
plot(bloom$Year,bloom$Avg_Hi, main="Average High Temperature by Year",xlab="Year", ylab="Degrees Celsius")
abline(lm(bloom$Avg_Hi~bloom$Year))
summary(lm(bloom$Avg_Hi~bloom$Year))
#plot(lm(bloom$Avg_Hi~bloom$Year)) #check residuals for heteroscedasticity and nonlinearity
```

```{r}
#This section runs linear regression on Avg Low temp by Year
#hist(bloom$Avg_Lo)
plot(bloom$Year,bloom$Avg_Lo, main="Average Low Temperature by Year",xlab="Year", ylab="Degrees Celsius")
abline(lm(bloom$Avg_Lo~bloom$Year))
summary(lm(bloom$Avg_Lo~bloom$Year))
#plot(lm(bloom$Avg_Lo~bloom$Year)) #check residuals for heteroscedasticity and nonlinearity
```

### not cumulative GDD
### This section runs linear regression on GDD by Year
```{r}
#hist(bloom$GDD)
#hist(bloom$GDD2) #if Tbase is -10
plot(bloom$Year,bloom$GDD, main="Average Growing Degree Days by Year",xlab="Year", ylab="Degrees Celsius")
abline(lm(bloom$GDD~bloom$Year))
summary(lm(bloom$GDD~bloom$Year))
#plot(lm(bloom$Avg_Lo~bloom$Year)) #check residuals for heteroscedasticity and nonlinearity


ggplot(bloom, aes(Year,GDD)) +
  geom_point()+
  stat_smooth() #uses loess()
```

```{r}

#This section runs linear regression on Precipitation by Year
hist(bloom$Raw_Precip)
plot(bloom$Year,bloom$Raw_Precip, main="Precipitation by Year",xlab="Year", ylab="Precipitation (mm)")
abline(lm(bloom$Raw_Precip~bloom$Year))
summary(lm(bloom$Raw_Precip~bloom$Year))
#plot(lm(bloom$Raw_Precip ~bloom$Year)) #check residuals for heteroscedasticity and nonlinearity
```



#Graph just the significant spp by year to see how many days earlier for these species that are showing a signal by year
#re-run code that runs regressions on each species to create table to subset from
#Limit all data to after 1949, 7220 obs. from table (7217 obs. from query in Access)
```{r}

bl50 <- subset(bloom, bloom$year_plant > 1949) #Our dataset already has this subset but re-running as a double check

nrow(bl50)
nrow(subset(bloom, bloom$year_plant > 1949))

str(bl50)
head(bl50)
```


```{r}
 names <- unique(bl50$Scientific.Name)		# Names includes synonyms, sspListID contains unique taxa
 namesID <- unique(bl50$sppListID)
length(namesID)	# 290
head(namesID)
names(bl50)

#There are multiple bloom dates per year for many species. Want to narrow down to one, earliest, per year
head(table(bl50$sppListID, bl50$Year))

collections<-data.frame(table(bl50$sppListID,bl50$Year))

ggplot(collections, aes(Var2, Freq))+
  geom_boxplot()+
  stat_smooth()+
  xlab("Year")+
  ylab("Collections species")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90, hjust=1))

# collections per month
collJ <- data.frame(table(bl50$sppListID, bl50$startDayOfYear))

ggplot(collJ, aes(Var2, Freq, group =1))+
  stat_smooth()+
  xlab("Julian date")+
  ylab("Collections per species")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90, hjust=1))

# collections getting 

```


Don't worry about bloom vs. bl50, they are identical... and other checks    
```{r}
#bl50 and bloom are the same
identical(bl50, bloom)

#There should be 290 species
length(unique(bl50$sppListID))

#Only need some columns of bloom:bl50 for minimum julian dates
names(bl50[,c(4:5,11,19:28)])
```


Pull out the earliest bloom date per year per species
## Break up the function into gathering the earliest bloom day per year seperate from regression per species
## Earliest bloom date held in min.dates
Aggregate instead of apply?
Much faster and more corrector!
```{r}

minJul.dates <- lapply(unique(bl50$sppListID), function(x){
  this.species <- bl50[bl50$sppListID == x,c(4:5,11,19:28)]
  these.rows <- aggregate(startDayOfYear~Year, this.species, FUN=min)
  unique(merge(this.species, these.rows, by = c("startDayOfYear","Year")))
  })

names(minJul.dates) <- unique(bl50$sppListID)

head(minJul.dates[[219]])

min.dates <- do.call(rbind,minJul.dates)

head(table(min.dates$Year, min.dates$sppListID))

```


#Check that there is only one minimum date per year
```{r}
minpyear <- table(min.dates$Year, min.dates$sppListID)
mpy.df <- data.frame(minpyear)
mpy.df[mpy.df$Freq > 1,] #when we have ealiestBloomDate in the lapply, that's not unique so it picks a few duplicates, now fixed
```

```{r}
library(psych)
#cor.plot(cor(min.dates[,-(1:4)]), numbers = TRUE)
# should keep Raw_Precip and Avg_Lo

cor(min.dates[,-(1:4)])

#getting a ggplot error of "Error in .Call.graphics(C_palette2, .Call(C_palette2, NULL)) : 
#  invalid graphics state"

#fix error by running dev.off()
# dev.off()

```


ggplot option after subset to earliest observation per year per species    
Avg high temperature    
```{r}
ggplot(min.dates, aes(Year,Avg_Hi))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Average High Temperature by Year")+
  xlab("Year")+
  ylab("Degrees Celsius")

summary(lm(Avg_Hi~Year, data=min.dates))

```

```{r}
ggplot(min.dates, aes(Avg_Hi, startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Average High Temperature") +
  xlab("Degrees Celsius")+
  ylab("Julian Date")

summary(lm(startDayOfYear~Avg_Hi, data=min.dates))
```

Average low temperature
```{r}
ggplot(min.dates, aes(Year,Avg_Lo))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Average Low Temperature by Year")+
  xlab("Year")+
  ylab("Degrees Celsius")

summary(lm(Avg_Lo~Year, data = min.dates))

```

```{r}
ggplot(min.dates, aes(Avg_Lo, startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Average Low Temperature") +
  xlab("Degrees Celsius")+
  ylab("Julian Date")

summary(lm(startDayOfYear~Avg_Lo, data = min.dates))
```



Average precipitation
```{r}
ggplot(min.dates, aes(Year,Raw_Precip))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Average Precipitation by Year")+
  xlab("Year")+
  ylab("Precipitation (mm)")


summary(lm(Raw_Precip~Year, data = min.dates))
```

```{r}
ggplot(min.dates, aes(Raw_Precip, startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Average Precipitation") +
  xlab("Precipitation (mm)")+
  ylab("Julian Date")

summary(lm(startDayOfYear~Raw_Precip, data = min.dates))
```

## not real GDD!
GDD    
```{r}
ggplot(min.dates, aes(Year,GDD))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Average Growing Degree Days by Year")+
  xlab("Year")+
  ylab("Growing Degree Day")


summary(lm(GDD~Year, data = min.dates))
```

```{r}
ggplot(min.dates, aes(GDD, startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Growing Degree Days") +
  xlab("GDD")+
  ylab("Julian Date")

summary(lm(startDayOfYear~GDD, data = min.dates))
```


### Minimum Julian bloom date by year
```{r}
ggplot(min.dates, aes(Year,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Year")+
  xlab("Year")+
  ylab("Julian Date")

summary(lm(min.dates$startDayOfYear~min.dates$Year))
max(min.dates$startDayOfYear)
```
    
    
    
Graph of all spp data by year using plot instead of ggplot

```{r}
#bloom<-read.csv(path.expand("P:/alpine-phenology/QR_final_R_plant_climate_data.csv"),        
#                header = TRUE, as.is=TRUE)   
hist(min.dates$startDayOfYear)#need to log transform data

plot(min.dates$Year,min.dates$startDayOfYear, main="First Collection Date for All 
Species",xlab="Year", ylab="Julian Date")
abline(lm(min.dates$startDayOfYear~min.dates$Year))
summary(lm(min.dates$startDayOfYear~min.dates$Year)) 
#both intercept and slope significant. If include all data, the date of first collection is statistically earlier by 0.062563/year, or 3.75 days earlier over the course of the study period. 
#plot(lm(bloom$startDayOfYear ~bloom$Year)) #check residuals for heteroscedasticity and nonlinearity
```



#Regression per species
#updated regression loop to match pulling minimum dates above    
The variable can be set to which variable you want to use. Choose from "Year", "Raw_Precip", "Avg_Hi", or "Avg_Lo"...    

```{r}
JulBloomDate <- function(variable){
  namesID <- unique(min.dates$sppListID)
  
  #Which column number has the variable of interest
  colnum <- match(variable,names(min.dates))
  
  JYC <- lapply(namesID, function(x){
    blnew <- min.dates[min.dates$sppListID == x,]
    foo1 <- summary(lm(blnew$startDayOfYear~blnew[,colnum]))
    jul.yr.coef <- data.frame(cbind(data.frame(t(foo1$coefficients[2,])),
                                    foo1$coefficients[1,1],
                                    t(foo1$fstatistic),
                                    Species = blnew$Scientific.Name[1],
                                    foo1$adj.r.squared))
    })
  
  JYC <- do.call(rbind, JYC)
  colnames(JYC) <- c("Slope", "StErr", "t.value", "p.value", "Intercept", 
                   "F Stat", "numDF","denDF", "Species", "adj.r.squ")
  JYC
}
```


# "Year" regression per species    

```{r}
yr.bl <- JulBloomDate("Year")

head(yr.bl)
#subset that signficantchange over time
yr.sig <- yr.bl[yr.bl$p.value < 0.05,]

#merge the significant species with the min.dates species
yr.sig.min <- merge(yr.sig, min.dates, by.x = "Species", by.y = "Scientific.Name")


ggplot(yr.sig.min, aes(Year,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Year")+
  xlab("Year")+
  ylab("Julian Date")

summary(lm(yr.sig.min$startDayOfYear~yr.sig.min$Year))

nrow(yr.sig) #53 species
nrow(yr.sig[yr.sig$Slope > 0,]) # no species have later bloom dates over time
nrow(yr.sig[yr.sig$adj.r.squ > 0.25,]) # 18 species have adj.r.squares over 0.25
yr.sig[yr.sig$adj.r.squ > 0.25,]


```



# "Raw_Precip" regression per species    

```{r}
precip.bl <- JulBloomDate("Raw_Precip")

head(precip.bl)
#subset that signficant change by precipitation
precip.sig <- precip.bl[precip.bl$p.value < 0.05,]

#merge the significant species with the min.dates species
precip.sig.min <- merge(precip.sig, min.dates, by.x = "Species", by.y = "Scientific.Name")


ggplot(precip.sig.min, aes(Raw_Precip,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by precipitation")+
  xlab("Precipitation (mm)")+
  ylab("Julian Date")

 summary(lm(precip.sig.min$startDayOfYear~precip.sig.min$Year))



```



# "GDD" regression per species    

```{r}
gdd.bl <- JulBloomDate("GDD")

head(gdd.bl)
#subset that signficant change by GDD
gdd.sig <- gdd.bl[gdd.bl$p.value < 0.05,]

#merge the significant species with the min.dates species
gdd.sig.min <- merge(gdd.sig, min.dates, by.x = "Species", by.y = "Scientific.Name")


ggplot(gdd.sig.min, aes(GDD,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by GDD")+
  xlab("GDD")+
  ylab("Julian Date")

 summary(lm(gdd.sig.min$startDayOfYear~gdd.sig.min$Year))


nrow(gdd.sig) #101 species
nrow(gdd.sig[gdd.sig$Slope > 0,]) # no species have later bloom dates as the GDD increases

```



########## Columns that you can use to compare to earliest bloom date (taken from "startDayOfYear")
              
#[11] "startDayOfYear"   [20] "Raw_Precip"                 
#[21] "Avg_Hi"                "Med_Hi"                      
#[23] "Avg_Lo"                "Med_Lo"                       
#[25] "Av_Temp"               "Med_Temp"       





# multiple linear regression

```{r}
  namesID <- unique(min.dates$sppListID)
  
  ml.JYC <- lapply(namesID, function(x){
    blnew <- min.dates[min.dates$sppListID == x,]
    foo1 <- summary(lm(startDayOfYear~Raw_Precip*Avg_Lo, data = blnew))
    jul.yr.coef <- data.frame(row.names(foo1$coefficients[-1,]),
                              foo1$coefficients[-1,],
                              foo1$coefficients[1,1],
                              t(foo1$fstatistic),
                              Species = blnew$Scientific.Name[1],
                              foo1$adj.r.squared)
    })
  
  ml.JYC <- do.call(rbind, ml.JYC)
  colnames(ml.JYC) <- c("Variables", "Slope", "StErr", "t.value", "p.value", "Intercept", 
                   "F Stat", "numDF","denDF", "Species", "adj.r.squ")
  head(ml.JYC)

sig.ml.JYC <- ml.JYC[ml.JYC$p.value < 0.05,]

# there are 19 species significant for Avg_Lo, 15 for Raw_Precip, and 14 for the interaction 
table(sig.ml.JYC$Variables)
```


# Species signficantly changing with the interaction of precipitation and low temperature
```{r}
sig.interaction.JYC <- ml.JYC[ml.JYC$p.value < 0.05 & ml.JYC$Variables == "Raw_Precip:Avg_Lo",]

sig.in.merge <- merge(sig.interaction.JYC, min.dates, by.x = "Species",
                      by.y = "Scientific.Name")


ggplot(sig.in.merge, aes(Year,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Year")+
  xlab("Year")+
  ylab("Julian Date")

 summary(lm(startDayOfYear~Year, data = sig.in.merge))


```

# Signficant interaction species     
Precip
```{r}
ggplot(sig.in.merge, aes(Raw_Precip,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Precipitation")+
  xlab("Precipitation (mm)")+
  ylab("Julian Date")

 summary(lm(startDayOfYear~Raw_Precip, data = sig.in.merge))

```


```{r}
ggplot(sig.in.merge, aes(Avg_Lo,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by Average Low Temperature")+
  xlab("Degrees Celcius")+
  ylab("Julian Date")

 summary(lm(startDayOfYear~Avg_Lo, data = sig.in.merge))

```


# extract climate data from .bil file time series    
<https://github.com/ropensci/prism>  
```{r, eval=FALSE}
# slice out only elevations above 10,500" (3200m)
latlongs <- read.delim(path.expand("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_QGIS Projects/Alpine Phenology Project/Sampling Points/LatLongabove3200m.txt"), sep=",")

ll <- mapply(FUN = function(x,y){
  c(x,y)
  },
  latlongs$Long,latlongs$Lat
  )

head(latlongs)
ll[,1]

prism_slice(ll[,1],ls_prism_data()[3,1]) # picks one max temp at one point
```

Into Climate_MEDL
```{r, eval=FALSE}
options(prism.path = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_QGIS Projects/Alpine Phenology Project/Climate_MEDL")
# takes a long long long time!!! Don't run again!
 get_prism_dailys(type="tmin", 
                 minDate = "1981-01-01", 
                 maxDate = "2011-12-31", keepZip = FALSE)


ls_prism_data()[3:13,]
ls_prism_data(absPath = TRUE)[3:13,] #to see path to where they are
ls_prism_data(name = TRUE)[3:13,] #details of the file
prism_image(ls_prism_data()[3,1])

ls_prism_data()[3,]


```

Into Climate_MEDL
```{r, eval=FALSE}
# error in download length != reported length
options(prism.path = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_QGIS Projects/Alpine Phenology Project/Climate_MEDL")

#get_prism_dailys(type="tmin", minDate = "1981-01-01", maxDate = "1981-02-01",
#                 keepZip = FALSE, mode="wb") #but I don't think mode is a thing...


# run 12/3/2015
get_prism_dailys(type="tmax", 
                 minDate = "1981-01-01", 
                 maxDate = "2011-12-31", keepZip = FALSE)
```

```{r}


```



Calculate average daily temperatures averaged over all locations above 3200m 10500"

```{r, eval=FALSE}
# combine by year
sliceyear <- grep("1981",ls_prism_data()[,1],value = TRUE)
sliceyear <- grep("tmin", sliceyear, value = TRUE) #only have tmin in this folder
str(sliceyear)
str(sliceyear[[1]][1])
p <- prism_slice(c(-106.688031183,40.9074207425), sliceyear)
prism_slice(ll[,1], sliceyear)
p$data

head(p$data)
head(ll[,1:10])
mean(p$data$data)

# going through first 10 days of 1981
ptest <- prism_slice(ll[,1],sliceyear[[1]][1:10]) 
head(ptest)
ptest$data

length(ls_prism_data()[,1]) #22,644

# test apply through a few points to get values for a mean
test1981 <- lapply(1:5, function(y){
  prism_slice(ll[,y], sliceyear[[1]][1:10])
  })

test1981[[1]]$data
# data frame each $data part and then bind them
test1981one <- do.call(rbind, lapply(test1981, function(x){
  do.call(data.frame, x$data)
  }))

#aggregate for average over all the locations
aggregate(data~date, data = test1981one, FUN = mean)

length(ll) #2529 above 3200m points
```

```{r testing, eval=FALSE}
# All the data held here
options(prism.path = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_QGIS Projects/Alpine Phenology Project/Climate_MEDL")

samplell <- sample(1:length(ll), 5)

# raster instead of slice!
avgTemps <- lapply(c(1:5,2000:2005), function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[1,2])
  r2points <- data.frame(rasterToPoints(rastertemps))  # pulls out data from raster
  r2p <- data.frame(r2points, colnames(r2points)[3])
  merge(r2points, latlongs[,3:4], by.x = c("x","y"), by.y = c("Long","Lat"))
})

# no points match, that's silly!
head(avgTemps[[1]])
```

Could limit days to months before latest of the earliest bloom days... 
```{r Temp data, eval=FALSE}
# raster and extract
maxs <- grep("tmax", ls_prism_data(absPath=TRUE)[,2])
mins <- grep("tmin", ls_prism_data(absPath=TRUE)[,2], value=TRUE)
#Jan - Sept, exclude Oct, Nov, and Dec. Latest first bloom is 269 = Sept 26
# exclude <- c("_[1-2][09][0-9][0-9][1][0]","_[1-2][09][0-9][0-9][1][1]","_[1-2][09][0-9][0-9][1][2]")
mins <- grep("_[1-2][09][0-9][0-9][0][1-9]", mins)

avgTemps.max <- lapply(maxs, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, latlongs[,c(4,3)]), date = ls_prism_data()[x,])
})

avgTemps.min <- lapply(mins, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, latlongs[,c(4,3)]), date = ls_prism_data()[x,])
})
```

```{r}
nrow(latlongs)
# they all got pulled out!!! yay!!
nrow(avgTemps.min[[1]])
nrow(avgTemps.max[[1]])


mins.avg <- do.call(rbind, avgTemps.min)
meanmins <- aggregate(data~date, data = mins.avg, mean)
maxs.avg <- do.call(rbind, avgTemps.max)
meanmaxs <- aggregate(data~date, data = maxs.avg, mean)

head(mins.avg)
head(as.Date(substr(meanmins$date, 25, 32), "%Y%m%d"))

# Make a day column
meanmins$day <- as.Date(substr(meanmins$date, 25, 32), "%Y%m%d")
meanmaxs$day <- as.Date(substr(meanmaxs$date, 25, 32), "%Y%m%d")

# make a julian day column
meanmins$Julian <- yday(meanmins$day)
meanmaxs$Julian <- yday(meanmaxs$day)

# merge min and max per day
temps <- merge(meanmins, meanmaxs, by = "day")  #.x min, .y max
tmps <- temps[,c(1,3:4,6)]
names(tmps) <- c("day","tmin","Julian","tmax")


#make a year column
tmps$Year <- as.numeric(format(temps$day, "%Y"))

head(tmps)

ggplot(tmps, aes(as.factor(Year)))+
  geom_boxplot(aes(y=tmin, colour = "red"))+
  geom_boxplot(aes(y=tmax, colour = "blue"))
```

Calculate cumulative HDD per year   
10 celsius is 50 fahrenheit   
5 celsius is 41  
Gordon and Bootsma (1993)   
```{r, eval=FALSE}
tbase <- 0

cumsum(((tmps$tmax-tmps$tmin)/2)-5)

HDD <- sapply(unique(tmps$Year), function(year){
  sapply(grep(year, tmps$Year), function(x){
    if( (((tmps$tmax[x] - tmps$tmin[x])/2)-tbase) < tbase){
      0
      } else {
        ((tmps$tmax[x] - tmps$tmin[x])/2)-tbase
        }
    })
  })

# now have daily HDD, need to cumsum each year, which is each list. 
head(HDD)

HDD.yrs <- lapply(HDD, function(x){
  cumsum(x)
})
  
head(HDD.yrs[[1]])
HDD.all <- do.call(c,HDD.yrs)
tmps$HDD <- HDD.all
head(tmps)
tail(tmps)
```

Combine HDD per Julian day for 1981 through 2011 with herbarium data
```{r}
# remove two attempts at not real HDD
min.dates <- min.dates[,-c((length(min.dates)-1):length(min.dates))]
head(min.dates)

#assign real HDD
min.dates <- merge(min.dates, tmps, 
                   by.x = c("Year","startDayOfYear"), by.y = c("Year","Julian"))

nrow(min.dates[complete.cases(min.dates),])
nrow(min.dates)

# HDD regression per species
hdd.bl <- JulBloomDate("HDD")
nrow(hdd.bl)
nrow(hdd.bl[complete.cases(hdd.bl),]) # 2 rows with missing data? why!?
hdd.bl[!complete.cases(hdd.bl),] # Kobresia schoenoides and Carex misandra

hdd.bl <- hdd.bl[complete.cases(hdd.bl),]
head(hdd.bl)
hdd.sig <- hdd.bl[hdd.bl$p.value < 0.05,]

#merge significant species iwth min.dates species
hdd.sig.min <- merge(hdd.sig, min.dates, by.x = "Species", by.y = "Scientific.Name")


nrow(hdd.sig) #278
nrow(hdd.bl) #288
hdd.bl[hdd.bl$p.value > 0.049,]



```

Average HDD? per year
```{r}
avg.HDD.yrs <- lapply(HDD.yrs, function(x){
  mean(x)
})

avg.yrs <- do.call(c, avg.HDD.yrs)
hddyrs <- cbind(Year=1981:2011, avgHDD = avg.yrs)


ggplot(data.frame(hddyrs), aes(Year, avgHDD))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Average cumulative HDD by Year")+
  xlab("Year")+
  ylab("Cumulative HDD")

```

```{r}
ggplot(hdd.sig.min, aes(HDD,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title= expression("HDD with Tbase = 0"~degree*C))+
  xlab("GDD")+
  ylab("Julian Date")

summary(lm(hdd.sig.min$startDayOfYear~hdd.sig.min$HDD))

```

HDD with Tbase = 0$^\circ$C
```{r}
ggplot(min.dates, aes(as.factor(Year), HDD))+
  geom_boxplot()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="HDD by Year")+
  xlab("Year")+
  ylab("HDD")

```

First bloom date by HDD for all species
```{r}
ggplot(min.dates, aes(HDD, startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="First bloom by HDD")+
  xlab("HDD")+
  ylab("First bloom date")

```

Heating degree days 

```{r}
ggplot(min.dates, aes(Year,HDD))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="HDD for each first bloom day")+
  xlab("Year")+
  ylab("HDD")
```

```{r}
ggplot(min.dates, aes(HDD, startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  labs(title="Earliest Bloom Date by HDD")+
  xlab("HDD")+
  ylab("Julian Date")
```


# Match herbarium data with NPN data
```{r}




```




## Can get worldclim data
```{r, eval=FALSE}
library(rgdal)

w <- getData('worldclim', var='tmin', res=0.5, lon=5, lat=45)


```
