---
title: "Final Methods and Results"
author: "Michelle DePrenger-Levin"
date: "March 3, 2016"
output: html_document
---

load libraries   
```{r}

library(RCurl)
library(ggplot2)
library(data.table)
library(raster)
require(rgdal)
library(devtools)
# devtools::install_github("ropensci/prism")
library(prism)
library(polynom)
library(MASS)
# install.packages("segmented")
library(segmented)
```

## load functions   
```{r}

JulBloomDate.subset <- function(variable,dataset){
  namesID <- unique(dataset$sppListID)
  colnum <- match(variable,names(dataset))
  JYC <- lapply(namesID, function(x){
    blnew <- dataset[dataset$sppListID == x,]
    foo1 <- summary(lm(blnew$startDayOfYear~blnew[,colnum]))
    jul.yr.coef <- data.frame(cbind(data.frame(t(foo1$coefficients[2,])),
                                    foo1$coefficients[1,1],
                                    t(foo1$fstatistic),
                                    Species = blnew$Scientific.Name[1],
                                    #Species = blnew$Species[1],
                                    foo1$adj.r.squared))
    })
  JYC <- do.call(rbind, JYC)
  colnames(JYC) <- c("Slope", "StErr", "t.value", "p.value", "Intercept", 
                   "F Stat", "numDF","denDF", "Species", "adj.r.squ")
  JYC
}

mergeMinGDD <- function(mindates,GDD){
  merge(mindates, GDD, by.x = c("Year","startDayOfYear"), by.y = c("Year","Julian"))
}

```


From Q:\Reserach\Plant Lists\
Data to subset
```{r}
colo <- read.csv("P:/hackathon/alpine-phenology/datasets/Complete COLO DB .csv") 
seinet <- read.csv(text = getURL("https://raw.githubusercontent.com/DenverBotanicGardens/alpine-phenology/master/CO_SEINet_3200_20160304.csv"))
```

Manipulate dates and names to be consistant before rbind databases
```{r}
colo$scientificName <- paste(colo$Genus,colo$Specific.Epithet,sep=" ")
names(colo[,c(1,40,13,31,32,21,22,23,24)])
colo$Reproductive.Status <- as.character(colo$Reproductive.Status)

# Select only reprodutive or post reproductive records
repro <- c("Flr","Bulblets","Seed","Fr")
colo.sm <- colo[colo$Reproductive.Status %in% unique(grep(paste(repro,collapse="|"),
                                                                colo$Reproductive.Status,
                                                                value=TRUE)),]

colo.sm <- colo[,c(1,40,12,13,31,32,21,22,23,24)]

#Year
colo.sm$Year <- (sub('.*(\\d{4}).*', '\\1', as.character(colo.sm$Date..dd.month.yyyy.)))

# Removing any with dates that aren't entered correctly or are missing year, not exact date
colo.sm <- colo.sm[grep("[[:digit:]]", colo.sm$Year),]
colo.sm <- colo.sm[colo.sm$Year>=1950,]

length(unique(colo.sm$scientificName)) #2999

# Only collections above or equal to 3200 meters, remove any without elevation
# only collections above or equal to 10,498.69 ft, remove any without elevation
# 15000 ft is 4572 m, remove anything over 4572 meters as a typo 
#  - highest in CO is Mt. Elbert at 4401.2 m
colo.sm$elevation <- (colo.sm$Elevation..feet.)/3.2808
colo.sm$elevation[!is.na(colo.sm$Elevation..meters.)] <-
  colo.sm$Elevation..meters.[!is.na(colo.sm$Elevation..meters.)]
colo.sm <- colo.sm[colo.sm$elevation>=3200,]
colo.sm <- colo.sm[!is.na(colo.sm$Elevation..meters.),]
colo.sm <- colo.sm[colo.sm$elevation < 4572,]

length(unique(colo.sm$scientificName)) #558

# add a julian day column as $startDayOfYear
colo.sm$Date..dd.month.yyyy. <- as.character(colo.sm$Date..dd.month.yyyy.)
wrongdates <- colo.sm$Date..dd.month.yyyy.[colo.sm$Date..dd.month.yyyy. %in% grep("-", colo.sm$Date..dd.month.yyyy., value=TRUE)]

table(wrongdates)

colo.sm$startDayOfYear<-as.numeric(format(
    as.Date(as.character(colo.sm$Date..dd.month.yyyy.), format="%d %B %Y")    ,"%j"))
```

SEInet  
pull out names to match colo.sm and have data needed to subset   
 [1] "Family.Name"          "scientificName"       "Country"              "State"                "Decimal.Latitude"      
 [6] "Decimal.Longitude"    "Elevation..feet."     "Elevation..meters."   "Date..dd.month.yyyy." "Reproductive.Status"   
[11] "Year"                 "elevation"            "startDayOfYear"   
```{r}
head(seinet[seinet$scientificName=="",])
seinet <- seinet[seinet$scientificName!="",]

head(seinet[,c(12:13,56,63:64,31,73,34)])
seinet$reproductiveCondition <- as.character(seinet$reproductiveCondition)

repro <- c("Fl","fr","Fert")
seinet.sm <- seinet[seinet$reproductiveCondition %in% unique(grep(paste(repro,collapse="|"),
                                                             seinet$reproductiveCondition,
                                                             value=TRUE)),]

seinet.sm <- seinet[,c(12:13,56,63:64,31,73,34)]

# get to 1950-2011
seinet.sm <- seinet.sm[seinet.sm$year >=1950,]
#seinet.sm <- seinet.sm[seinet.sm$year <= 2011,]
seinet.sm <- seinet.sm[!is.na(seinet.sm$year),]

# Elevation over 3200 meters - was gathered from SEINet with that limitation. 
nrow(seinet.sm[is.na(seinet.sm$minimumElevationInMeters),]) #498
seinet.sm <- seinet.sm[!is.na(seinet.sm$minimumElevationInMeters),]
identical(seinet.sm,seinet.sm[seinet.sm$minimumElevationInMeters >=3200,])
```

Combine datasets to create bloom2   
```{r}
names <- names(colo.sm[,c(1:2,5:6,11:13)])
#names(seinet.sm[,c(1:4,7,5,8)])
colo.bind <- colo.sm[,c(1:2,5:6,11:13)]
seinet.bind <- seinet.sm[,c(1:2,4:8)]
names(seinet.bind) <- names

i<-sapply(colo.bind, is.factor)
colo.bind[i]<-lapply(colo.bind[i], as.character)
colo.bind$Year <- as.numeric(colo.bind$Year)
i<-sapply(seinet.bind, is.factor)
seinet.bind[i]<-lapply(seinet.bind[i],as.character)
i<-sapply(seinet.bind, is.integer)
seinet.bind[i]<-lapply(seinet.bind[i],as.numeric)

bloom2 <- rbind(seinet.bind,colo.bind) 
barplot(table(bloom2$Year))

length(unique(bloom2$scientificName)) #2303
unique(bloom2$Family.Name)[1:100]
proper <- function(s) sub("(.)", ("\\U\\1"), tolower(s), pe=TRUE)
bloom2$Family.Name <- proper(bloom2$Family.Name)
```

Fix spelling errors   
<http://cumuseum-archive.colorado.edu/Research/Botany/Databases/vascular_plants.pdf>
```{r}
bloom2 <- as.data.frame(sapply(bloom2, gsub, 
                               pattern = "caerulea",
                               replacement = "coerulea"))

bloom2 <- as.data.frame(sapply(bloom2, gsub,
                               pattern = "Eritrichum",
                               replacement = "Eritrichium"))

str(bloom2)
tochar <- c(1:2)
bloom2[,tochar] <- sapply(bloom2[,tochar], function(ch){
  as.character(ch)
})
rm(tochar)

tonumeric <- c(3:length(bloom2))
bloom2[,tonumeric] <- sapply(bloom2[,tonumeric], function(s){
  as.numeric(as.character(s))
})
```


#Check for synonyms   
```{r}
synonyms <- read.csv(text = getURL("https://raw.githubusercontent.com/DenverBotanicGardens/alpine-phenology/master/datasets/synonym_list_byhand2.txt"))

length(grep("ssp.",bloom2$scientificName)) #1577 (75)
length(grep("subsp.",bloom2$scientificName)) #1502 (0)
length(grep("ssp.",synonyms$SYNONYMS)) #6400 -> 6426
length(grep("subsp.",synonyms$SYNONYMS)) #27 -> 0

#use of ssp and subsp
synonyms$SYNONYMS <- gsub(pattern = "subsp.", replacement = "ssp.", synonyms$SYNONYMS)
bloom2$scientificName <- gsub(pattern = "subsp.", replacement = "ssp.", bloom2$scientificName)

# names before the merge from our dataset
namebloom2 <- unique(bloom2$scientificName)
length(namebloom2) #2284

# ID links synonyms
# to not lose ones from bloom2 that have no match in synonyms, add all.x = TRUE (left outer join)
bloom.lo <- merge(bloom2, synonyms[,-3], by.x = "scientificName", by.y = "SYNONYMS", all.x = TRUE)
bloom.o <- merge(bloom2, synonyms[,-3], by.x = "scientificName", by.y = "SYNONYMS")

length(unique(synonyms$SYNONYMS)) #91,084 names (91094 after adding 12 by hand - 2 were repeated?)
length(unique(synonyms$ID)) #48,141 unique taxa
length(unique(bloom.lo$scientificName)) #2284
length(unique(bloom.o$scientificName)) #1986 -> 1996 after adding ones by hand
length(unique(bloom2$scientificName)) #(2284) 2299 different names in bloom2 before merge - then 1986 after; 2284

#which names are in bloom2 before merge (namebloom2) vs. after merge (bloom2$scientificName)
badnames.o <- setdiff(namebloom2,bloom.o$scientificName)
head(badnames.o)
length(badnames.o) #288

badnames2 <- namebloom2[!(namebloom2 %in% bloom2$scientificName)]
length(badnames2) #471
badnames3 <- bloom2$scientificName[!(bloom2$scientificName %in% namebloom2)]
length(badnames3) #0
write.table(badnames.o,"clipboard", sep="\t", row.names=FALSE)

#Use these 'badnames' which are names in bloom2 not found in the USDA synonym list to correct errors
# some are crosses, some misspellings. 
# csv has $Wrong: list of bloom2 names with no UDSA match 
#         $Replacement: corrected or NA if it should just be removed/ignored
# Should just remove ssp. and var. to lump. 
fixbloom2 <- read.csv(path.expand("Q:/Research/Projects/alpine-phenology/SpellingErrorCorrections.csv"),
                      na.strings = c("NA", "", " "), colClasses = "character")

#try again, first replace names with corrections, then do the merge for left outer again)
fixbloom2.all <- fixbloom2[complete.cases(fixbloom2[,2]),-3]

#maybe try a merge - make the new column scientificName be same if no matching, new if matching

bloomfoo <- bloom2

#where is there a match for each?
matches <- apply(fixbloom2.all,1,function(checknames){
  grep(checknames[1],bloomfoo$scientificName)
})
allmat <- do.call(c,matches)
length(allmat) #679 switches
```


```{r}
multisub <- function(pattern,replacement,x){
  result <- x
  for(i in 1:length(pattern)){
    result <- gsub(pattern[i], replacement[i],result)
  }
  result
}

bloomfoo$scientificName <- multisub(fixbloom2.all$Wrong,fixbloom2.all$Replacement,bloomfoo$scientificName)

str(bloomfoo)
```

Doesn't seem to be changing the data frame types
```{r}
tochar <- 1
bloomfoo[,tochar] <- sapply(bloomfoo[,tochar], function(ch){
  as.character(ch)
})
rm(tochar)

tonumeric <- c(3:length(bloomfoo))
bloomfoo[,tonumeric] <- sapply(bloomfoo[,tonumeric], function(s){
  as.numeric(as.character(s))
})

```

Did it work?
Sure looks like it did!!
```{r}
bloomfoo[bloomfoo$scientificName %in% "Aconitum columbianum",]
bloomfoo[bloomfoo$scientificName %in% "Aconitum columbianum Nutt.",]
bloom2[bloom2$scientificName %in% "Aconitum columbianum Nutt.",]

bloomfoo[bloomfoo$scientificName %in% "Agoseris aurantiaca",]
bloomfoo[bloomfoo$scientificName %in% "Agoseris aurantiaca var.",]
bloom2[bloom2$scientificName %in% "Agoseris aurantiaca var.",]

# Ok, do it!
bloom2 <- bloomfoo

# Now merge with synonyms
bloom.temp <- merge(bloom2, synonyms[,-3], by.x = "scientificName", by.y = "SYNONYMS")

str(bloom.temp)
bloom2 <- bloom.temp

```

How many species have two species ID numbers?
```{r}
foo <- data.frame(table(bloom2$scientificName,bloom2$ID))
foo[foo$Freq > 0,]
sort(table(foo$Var1[foo$Freq>0]))
```
Agrostis variabilis    
Anemone multifida    
Antennaria parvifolia   
Antennaria umbrinella    
Bromus anomalus    
Carex atrata    
Carex hallii    
Dasiphora fruticosa    
Eleocharis  
Erigeron acris     
Festuca ovina    
Galium boreale     
Linum lewisii   
Oxytropis deflexa       
Paronychia pulvinata    
Rorippa teres      
Salix anglorum     
Salix arctica      
Senecio cymbalarioides     


After combining datasets, find species with at least 10 years of data
```{r}
# all important data is there
bloom2[bloom2$scientificName=="",]
bloom2[bloom2$Year=="",]
bloom2[bloom2$ID=="",]

#how many species in dataset
length(unique(bloom2$scientificName)) #2294 -> 1828 -> 2207 -> 2018
length(unique(bloom2$ID)) # 1661 grouped by synonyms

# how many years collected per species
sp.yr <- as.data.frame(table(bloom2$Year,bloom2$ID))
nrow(sp.yr[sp.yr$Var2=="",])

nrow(sp.yr[sp.yr$Freq>0,]) #16685 -> 16666 -> 14498 after synonyms -> 15121
mean(sp.yr$Freq)
mean(sp.yr$Freq[sp.yr$Freq > 0])
max(sp.yr$Freq[sp.yr$Freq>0]) #80
head(sp.yr)

# How many vouchers per species per year
colpyear <- data.frame(table(bloom2$ID,bloom2$Year))
colpyear.2 <- do.call(rbind,lapply(split(colpyear,as.factor(colpyear$Var2)),
                                   function(x) x[which.max(x$Freq),]))

colpyear.2[with(colpyear.2, order(colpyear.2$Freq)),]

# change all values above 0 to 1 from number of collections for that species a year
# finding which years have any data, not the minimum date
sp.yr$Freq <- sapply(sp.yr$Freq,function(x){
  if(x>0) {1}
  else 0
  }, USE.NAMES = FALSE)

sp.yr.sp <- sp.yr[sp.yr$Freq>0,]
bysp <- split(sp.yr.sp, sp.yr.sp$Var2, drop=TRUE)
names(bysp[1])

spyears <- do.call(rbind,lapply(bysp, function(x){
  data.frame(DataYrs = sum(x$Freq),ID = unique(x$Var2))
}))

spyears[with(spyears, order(spyears$DataYrs)),]
spyears[which.max(spyears$DataYrs),] # 52
nrow(spyears[spyears$DataYrs>=10,]) #544 -> 479 -> 498
mean(spyears$DataYrs[spyears$DataYrs>=10]) #23.65553

sp.10plus <- spyears[spyears$DataYrs>=10,]
```


Pull only the species that have 10 or more years of data for analyses   
Not earliest but does tell me if one per species per year to tell how many species have at least 10 years a year   
```{r}
tenyearplus <- merge(bloom2,sp.10plus,by = "ID")
length(unique(sp.10plus$ID)) # 498
length(unique(tenyearplus$scientificName)) #545 -> 544 -> 449 -> 741
length(unique(tenyearplus$ID)) # 498

synlist <- lapply(split(tenyearplus, tenyearplus$ID), function(x){
  data.frame(Names = unique(x$scientificName), ID = unique(x$ID))
})

synlist <- do.call(rbind,synlist)
synlist[1:30,]
```



Pull out earliest collection   
```{r}
x<- unique(tenyearplus$ID)[100]

minJul.dates.2 <- lapply(unique(tenyearplus$ID), function(x){
  this.species <- tenyearplus[tenyearplus$ID == x,]
  these.rows <- aggregate(startDayOfYear~Year, this.species, FUN=min)
  merge(unique(this.species[,c("ID","Year","startDayOfYear","DataYrs")]), these.rows,
        by=c("startDayOfYear","Year"))
  })

min1950_2 <- do.call(rbind,minJul.dates.2)
head(min1950_2)

length(unique(min1950_2$ID)) #498 ! Yay!
length(unique(min1950_2$Family.Name)) #51
length(unique(min1950_2$ID))


```

```{r}
# check that none have more than one a year
singlefoo <- data.frame(table(min1950_2$Year,min1950_2$ID))
nrow(singlefoo[singlefoo$Freq> 1,]) #154? -> 0 Yay!

# add back the first species name (not the synonyms) for each data ID
onename <- bloom2[!duplicated(bloom2$ID),]

foo <- data.frame(table(min1950_2$scientificName,min1950_2$ID))
foo[foo$Freq>1,]

foo <- data.frame(table(onename$scientificName,onename$ID))
table(foo$Var1,foo$var2)

min1950_2.foo <- merge(min1950_2, onename[,c("scientificName","Family.Name","ID")], by="ID")
head(min1950_2.foo)
length(unique(min1950_2.foo$ID)) #498! Yay!
min1950_2 <- min1950_2.foo

#Check that there is only one species collection date per year
singlefoo <- data.frame(table(min1950_2$Year,min1950_2$ID))
nrow(singlefoo[singlefoo$Freq> 1,]) #11? -> 1?? -> 0

length(unique(min1950_2$ID)) #498 species
mean(min1950_2$DataYrs[!duplicated(min1950_2$ID)]) #23.665 years

nrow(min1950_2) #11673
write.table(min1950_2, file="P:/hackathon/alpine-phenology/datasets/min1950_2.csv", sep=",")
```

min1950_2 from away
```{r}
min1950_2 <- read.csv(text = getURL("https://raw.githubusercontent.com/DenverBotanicGardens/alpine-phenology/master/datasets/min1950_2.csv"))


```


Of the first bloom data of species that matched sampling critera, the hightest diversity of species collections were made in the 1990s and early 2000s with a peak of 449 species collected in 1998. Average over the study period was 211 alpine/subalpine species collected per year. 
```{r}
earliestcollection <- data.frame(colSums(table(min1950_2$scientificName, #should be only one name now
                                                min1950_2$Year)))

earcol <- data.frame(Year = row.names(earliestcollection),
                     Collections = earliestcollection)
colnames(earcol)[2] <- "Collections"
earcol[,1] <- as.numeric(levels(earcol[,1]))
earcol[which.max(earcol$Collections),]
mean(earcol$Collections) #191.36

```
        Year    Collections    
1998    1998            416    


#Figure 1A
```{r}
ggplot(earcol, aes(Year, Collections))+
  geom_bar(stat="identity")+
  xlab("Year")+
  ylab("Number of Study Species")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  scale_x_continuous(limits = c(1950,2011), breaks = seq(1950,2010, by=10))

```

 
```{r}
allcollections <- data.frame(colSums(table(bloom2$scientificName,bloom2$Year)))
allcol <- data.frame(Year = row.names(allcollections),
                     allcollections)
colnames(allcol)[2] <- "collections"

as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
allcol[,1] <- as.numeric.factor(allcol[,1])
```

#Figure 1B  
```{r}
ggplot(allcol,  aes(x = Year, y = collections))+
  geom_bar(stat="identity")+
  xlab("Year")+
  ylab("Total Collections")+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90, hjust=1))+
  scale_x_continuous(limits = c(1950,2011), breaks = seq(1950,2011, by=10))+
  scale_y_continuous(limits = c(0,4700),breaks = seq(0,4700, by=500))
```

The day each year with the most collections has on average been Julian day 207 and the day with the most collections has not moved significantly over time from 1950 to 2011. The earliest peak collection day was Julian day 175 (June 30th) in 2007 and the latest peak collection day was Julian day 238 in 1957 (August 29th)   
```{r}
collJ.sp <- data.frame(table(bloom2$Year,bloom2$startDayOfYear))

collJ.max <- lapply(split(collJ.sp, collJ.sp$Var1), function(x) x[which.max(x$Freq),])
collmax <- do.call(rbind,collJ.max)
collmax[,1:2]<-sapply(collmax[,1:2], function(x) as.numeric(as.character(x)))

mean(collmax$Var2)
collmax[which.min(collmax$Var2),]
collmax[which.max(collmax$Var2),]
summary(lm(Var2~Var1, data=collmax))

```


Average high, low, and precip per year  
Take annual climate data from previous dataset
```{r}
head(bloom)
head(bloom[,c(19,20,21,23)])
head(min1950_2)
min1950_2 <- merge(min1950_2,unique(bloom[,c(19,20,21,23)]), by="Year")

write.table(min1950_2, file="P:/hackathon/alpine-phenology/datasets/min1950_2.csv", sep=",")

unbloom <- unique(bloom[,c(19,20,21,23)])
```

Correlations of climate
```{r}
cor(unbloom[,-1])
cor.test(unbloom[,2],unbloom[,3])
cor.test(unbloom[,2],unbloom[,4])
cor.test(unbloom[,3],unbloom[,4])

```

#Figure 2 Year x Bloom date
```{r}
ggplot(min1950_2, aes(Year, startDayOfYear)) +
  geom_point(shape=1)+
  stat_smooth(method="lm")+
  theme_bw()+
  xlab("Year")+
  ylab("Ordinal date")+
  scale_x_continuous(breaks = seq(1950,2011,10))  

summary(lm(startDayOfYear~Year, data=min1950_2))

```

#Figure 2A  
Using just the average temp per year, not repeated for everytime we have specimen data in a year
```{r}
ggplot(unbloom, aes(Year,Avg_Hi))+
  geom_point(shape=1)+
  stat_smooth(method="lm")+
#  stat_smooth(geom = "smooth")+
  theme_bw()+
#  labs(title="Average High Temperature by Year")+
  xlab("Year")+
  ylab(expression(~degree~C)) + #ylab(expression("Average high temperature"~degree~C))
  scale_x_continuous(breaks = seq(1950,2011,10))

summary(lm(Avg_Hi~Year, data=unbloom))
```


#Figure 2B
```{r}
ggplot(unbloom, aes(Year,Avg_Lo))+
  geom_point(shape=1)+
  stat_smooth(method="lm")+
  theme_bw()+
#  labs(title="Average Low Temperature by Year")+
  xlab("Year")+
  ylab(~degree~C)+
  scale_x_continuous(breaks = seq(1950,2011,10))

summary(lm(Avg_Lo~Year, data = unbloom))

```

#Figure 2C   
```{r}
ggplot(unbloom, aes(Year,Raw_Precip))+
  geom_point(shape=1)+
  stat_smooth(method="lm")+
  theme_bw()+
#  labs(title="Average Precipitation by Year")+
  xlab("Year")+
  ylab("Precipitation (mm)")+
  scale_x_continuous(breaks = seq(1950,2011,10))


summary(lm(Raw_Precip~Year, data = unbloom))

```


#Figure 3A  
```{r}
ggplot(min1950_2, aes(Avg_Hi, startDayOfYear))+
  geom_point(shape=1)+
  stat_smooth(method="lm")+
  theme_bw()+
#  labs(title="Earliest Bloom Date by Average High Temperature") +
  xlab(expression(~degree~C))+
  ylab("Ordinal Date")

summary(lm(startDayOfYear~Avg_Hi, data=min1950_2))

```


#Figure 3B
```{r}
ggplot(min1950_2, aes(Avg_Lo, startDayOfYear))+
  geom_point(shape=1)+
  stat_smooth(method="lm")+
  theme_bw()+
#  labs(title="Earliest Bloom Date by Average Low Temperature") +
  xlab(expression(~degree~C))+
  ylab("Ordinal Date")

summary(lm(startDayOfYear~Avg_Lo, data = min1950_2))
```

#Figure 3C
```{r}
ggplot(min1950_2, aes(Raw_Precip, startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
#  labs(title="Earliest Bloom Date by Average Precipitation") +
  xlab("Precipitation (mm)")+
  ylab("Ordinal Date")

summary(lm(startDayOfYear~Raw_Precip, data = min1950_2))

```

Growing Degree Days   
```{r, eval=FALSE}
options(prism.path = "Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_QGIS Projects/Alpine Phenology Project/Climate_MEDL")

latlongs <- read.delim(path.expand("Q:/Research/All_Projects_by_Species/aa_Shapefiles_Maps/aa_QGIS Projects/Alpine Phenology Project/Sampling Points/LatLongabove3200m.txt"), sep=",")

maxs <- grep("tmax", ls_prism_data(absPath=TRUE)[,2])
mins <- grep("tmin", ls_prism_data(absPath = TRUE)[,2])
# mins <- grep("_[1-2][09][0-9][0-9][0][1-9]", mins)

# Takes an hour to run
avgTemps.max <- lapply(maxs, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, latlongs[,c(4,3)]), date = ls_prism_data()[x,])
})

avgTemps.min <- lapply(mins, function(x){
  rastertemps <- raster(ls_prism_data(absPath=TRUE)[x,2])
  data.frame(data = extract(rastertemps, latlongs[,c(4,3)]), date = ls_prism_data()[x,])
})


mins.avg <- do.call(rbind, avgTemps.min)
meanmins <- aggregate(data~date, data = mins.avg, mean)
maxs.avg <- do.call(rbind, avgTemps.max)
meanmaxs <- aggregate(data~date, data = maxs.avg, mean)
# Make a day column
meanmins$day <- as.Date(substr(meanmins$date, 25, 32), "%Y%m%d")
meanmaxs$day <- as.Date(substr(meanmaxs$date, 25, 32), "%Y%m%d")

# make a julian day column
meanmins$Julian <- yday(meanmins$day)
meanmaxs$Julian <- yday(meanmaxs$day)

# merge min and max per day
temps <- merge(meanmins, meanmaxs, by = "day")  #.x min, .y max
tmps <- temps[,c(1,3:4,6)]
names(tmps) <- c("day","tmin","Julian","tmax")


#make a year column
tmps$Year <- as.numeric(format(temps$day, "%Y"))

```


Calculate Growing Degree Days
```{r, eval=FALSE}
tbase <- 0

#method 1 from McMaster and Wilhelm 1997
HDD <- sapply(unique(tmps$Year), function(year){
  sapply(grep(year, tmps$Year), function(x){
    if( (((tmps$tmax[x] + tmps$tmin[x])/2)-tbase) < tbase){
      0
      } else {
        ((tmps$tmax[x] + tmps$tmin[x])/2)-tbase
        }
    })
  })

HDD.yrs <- lapply(HDD, function(x){
  cumsum(x)
})

HDD.all <- do.call(c,HDD.yrs)
tmps$HDD <- HDD.all

nrow(min1950_2[min1950_2$Year > 1980,]) # 7737 -> 7730 -> 7447

#min.dates <- merge(min.dates, tmps, 
#                   by.x = c("Year","startDayOfYear"), by.y = c("Year","Julian"))

min1950GDD_2 <- mergeMinGDD(min1950_2,tmps)

length(unique(min1950GDD_2$day)) #1152 -> 1149
length(unique(min1950GDD_2$scientificName)) #544 -> 493

#Years of data for GDD years per species
gdddatayrs <- as.data.frame(table(min1950GDD_2$Year,min1950GDD_2$scientificName))
yrspsp <- aggregate(gdddatayrs[,3], by = list(gdddatayrs$Var2), FUN=sum)
nrow(yrspsp[yrspsp$x>=10,]) #407 -> 376


### Reduce min1950GDD_2 to only species where 10 or more years of data
min1950GDD_2_10 <- merge(min1950GDD_2,yrspsp[yrspsp$x>=10,],
                         by.x="scientificName",by.y="Group.1")
length(unique(min1950GDD_2_10$scientificName)) #545 -> 376

```

correlations with GDD and climate factors
```{r}
.unbloomGDD <- merge(unbloom,hddyrs,by="Year")
cor(unbloomGDD[,-1])
cor.test(unbloomGDD$avgHDD,unbloomGDD$Raw_Precip)
cor.test(unbloomGDD$avgHDD,unbloomGDD$Avg_Hi)
cor.test(unbloomGDD$avgHDD,unbloomGDD$Avg_Lo)

```

#Figure 4A   
Average GDD for the ealiest collection date did not change significantly over the study years 
```{r}
ggplot(min1950GDD_2_10, aes(Year,HDD))+
  geom_point()+
  stat_smooth(method = "lm")+
  theme_bw()+
#  labs(title= expression("Cumulative GDD (Tbase = 0"~degree*C~")"))+
  xlab("Year")+
  ylab(expression(paste("GDD at collection date (T"[base] == 0, degree*C,")")))

summary(lm(HDD~Year, data=min1950GDD_2_10))
```

Growing Degree Days over time
```{r}

ggplot(unbloomGDD, aes(Year,avgHDD))+
  geom_point()+
  stat_smooth(method = "lm")+
  theme_bw()+
#  labs(title= expression("Cumulative GDD (Tbase = 0"~degree*C~")"))+
  xlab("Year")+
  ylab("cummulative GDD at collection date")

summary(lm(avgHDD~Year, data=unbloomGDD))
```


#Figure 4B
```{r}
totalGDD <- lapply(HDD.yrs, function(x){
  max(x)
})
totalGDDyr <- data.frame(Year=1981:2011,MaxGDD = do.call(rbind,totalGDD))
head(totalGDDyr)

ggplot(totalGDDyr, aes(Year,MaxGDD))+
  geom_point()+
  stat_smooth(method = "lm")+
  theme_bw()+
#  labs(title= expression("Maximum cumulative GDD (Tbase = 0"~degree*C~")"))+
  xlab("Year")+
  ylab(expression(paste("GDD at ordinal date 273 (T"[base] == 0, degree*C,")")))
summary(lm(MaxGDD~Year,data=totalGDDyr))

```


#Figure 4C
```{r}
#meltd is from Results figures and tables.Rmd
fit1 <- lapply(unique(meltd$Year), function(x){
  stepAIC(lm(GDD~Day+I(Day^2), data=meltd[meltd$Year==x,]),
          scope = lm(GDD~1, data=meltd[meltd$Year==x,]),
          direction = "backward")
})

poly.p <- lapply(fit1, function(fx){
  polynomial(coef(fx))
})
sday <- 1
eday <- 273

secant.poly <- lapply(poly.p,function(x){
  ((as.function(x)(eday)+1000) - (as.function(x)(sday)+1000))/(eday-sday)
})
secant.poly.all <- data.frame(Year=1981:2011, 
                              Secant= do.call(rbind,secant.poly))

ggplot(secant.poly.all, aes(Year,Secant))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
  ylab("Average rate of change in GDD (day 1 to 273)")

summary(lm(Secant~Year, data=secant.poly.all))

```

#Figure 4D 
```{r}
min1950GDDsecant2 <- merge(min1950GDD_2_10, secant.poly.all,by="Year")
head(min1950GDDsecant2)

ggplot(min1950GDDsecant2, aes(Secant,startDayOfYear))+
  geom_point()+
  stat_smooth(method="lm")+
  theme_bw()+
#  labs(title="Earliest Bloom Date by average rate of GDD accumulation") +
  xlab("Average rate of GDD accumulation")+
  ylab("Ordinal Date")

summary(lm(startDayOfYear~Secant, data=min1950GDDsecant2))

```

```{r, eval=FALSE}
ggplot(min1950GDD_2_10, aes(HDD,startDayOfYear))+
  geom_point()+
  stat_smooth(method = "lm")+
  theme_bw()+
  labs(title= expression("Cumulative GDD (Tbase = 0"~degree*C~")"))+
  xlab("GDD")+
  ylab("Ordinal Date")

summary(lm(startDayOfYear~HDD, data=min1950GDD_2_10))

```


NPN species   
(from Climate)    
```{r}
npn.all <- read.csv(path.expand("Q:/Research/Projects/alpine-phenology/NPN data/2015_species_list_complete.csv"))
npn.all$SciName <- apply(npn.all[,c("Genus","Species")], 1,
                         function(x) paste(x[1],x[2], collapse=" "))

#although lots of these might not have 10 years    
#Match names - synonyms and all
npn.big <- intersect(npn.all$SciName, bloom2$scientificName) 
length(npn.big) #27 -> 77

# find how many unique IDs match
npn.IDs <- bloom2$ID[bloom2$scientificName %in% npn.big]
length(unique(npn.IDs)) #78

# How many IDs that match have collections within 1950 to 2011
npn.foo <- min1950_2[min1950_2$ID %in% npn.IDs,]
nrow(npn.foo)
length(unique(npn.foo$ID)) #29
sort(unique(npn.foo$scientificName))
```

Clunky but need to change the function
```{r}
JulBloomDate.min1950 <- function(variable,dataset){
  namesID <- unique(dataset$ID)
  colnum <- match(variable,names(dataset))
  JYC <- lapply(namesID, function(x){
    blnew <- dataset[dataset$ID == x,]
    foo1 <- summary(lm(blnew$startDayOfYear~blnew[,colnum]))
    jul.yr.coef <- data.frame(cbind(data.frame(t(foo1$coefficients[2,])),
                                    foo1$coefficients[1,1],
                                    t(foo1$fstatistic),
                                    blnew$scientificName[1],
                                    blnew$Family.Name[1],
                                    foo1$adj.r.squared))
    })
  JYC <- do.call(rbind, JYC)
  colnames(JYC) <- c("Slope", "StErr", "t.value", "p.value", "Intercept", 
                   "F Stat", "numDF","denDF", "Species", "Fam", "adj.r.squ")
  JYC
}

```


Phenology summary tables    
```{r}
yr.1950_2 <- JulBloomDate.min1950("Year",min1950_2)

nrow(yr.1950_2)

write.table(yr.1950_2, file = "Q:/Research/Projects/alpine-phenology/phensum_yr.csv", sep=",")

nrow(yr.1950_2) #545 -> 498
nrow(yr.1950_2[yr.1950_2$p.value>=0.05,]) #475 -> 440
nrow(yr.1950_2[yr.1950_2$p.value<0.05,]) #70 -> 58

nrow(yr.1950_2[yr.1950_2$p.value<0.05,])/nrow(yr.1950_2) #0.1164

#earlier
nrow(yr.1950_2[yr.1950_2$p.value<0.05 & yr.1950_2$Slope<0,]) #65 -> 57
mean(yr.1950_2$Slope[yr.1950_2$p.value<0.05 & yr.1950_2$Slope<0]) 
mean(yr.1950_2$adj.r.squ[yr.1950_2$p.value<0.05 & yr.1950_2$Slope<0])

#later
nrow(yr.1950_2[yr.1950_2$p.value<0.05 & yr.1950_2$Slope>=0,]) #5 -> 1
mean(yr.1950_2$Slope[yr.1950_2$p.value<0.05 & yr.1950_2$Slope>=0])
mean(yr.1950_2$adj.r.squ[yr.1950_2$p.value<0.05 & yr.1950_2$Slope>=0])

```


###High Temperature     
```{r}
high.1950_2 <- JulBloomDate.min1950("Avg_Hi", min1950_2)

nrow(high.1950_2) #545 -> 498

write.table(high.1950_2, file = "Q:/Research/Projects/alpine-phenology/phen_summary_high_2.csv", sep=",")

nrow(high.1950_2[high.1950_2$p.value<0.05,])
nrow(high.1950_2[high.1950_2$p.value<0.05,])/nrow(high.1950_2)

#earlier
nrow(high.1950_2[high.1950_2$p.value<0.05 & high.1950_2$Slope<0,]) 
mean(high.1950_2$Slope[high.1950_2$p.value<0.05 & high.1950_2$Slope<0]) 
mean(high.1950_2$adj.r.squ[high.1950_2$p.value<0.05 & high.1950_2$Slope<0])

#later
nrow(high.1950_2[high.1950_2$p.value<0.05 & high.1950_2$Slope>=0,]) 
mean(high.1950_2$Slope[high.1950_2$p.value<0.05 & high.1950_2$Slope>=0])
mean(high.1950_2$adj.r.squ[high.1950_2$p.value<0.05 & high.1950_2$Slope>=0])

nrow(high.1950_2[high.1950_2$p.value<0.05,])/nrow(high.1950_2)
nrow(high.1950_2[high.1950_2$p.value>=0.05,])/nrow(high.1950_2)

```

```{r}
low.1950_2 <- JulBloomDate.min1950("Avg_Lo", min1950_2)

nrow(low.1950_2) #545  -> 498

write.table(low.1950_2, file = "Q:/Research/Projects/alpine-phenology/phen_sum_low_2.csv", sep=",")

nrow(low.1950_2[low.1950_2$p.value<0.05,]) #79
nrow(low.1950_2[low.1950_2$p.value<0.05,])/nrow(low.1950_2)

#earlier
nrow(low.1950_2[low.1950_2$p.value<0.05 & low.1950_2$Slope<0,]) #75 -> 68
mean(low.1950_2$Slope[low.1950_2$p.value<0.05 & low.1950_2$Slope<0]) 
mean(low.1950_2$adj.r.squ[low.1950_2$p.value<0.05 & low.1950_2$Slope<0])

#later
nrow(low.1950_2[low.1950_2$p.value<0.05 & low.1950_2$Slope>=0,]) #4 -> 3
mean(low.1950_2$Slope[low.1950_2$p.value<0.05 & low.1950_2$Slope>=0])
mean(low.1950_2$adj.r.squ[low.1950_2$p.value<0.05 & low.1950_2$Slope>=0])
```

### Precipitation
```{r}
precip.1950_2 <- JulBloomDate.min1950("Raw_Precip", min1950_2)

nrow(precip.1950_2) #545  -> 498

write.table(precip.1950_2, file = "Q:/Research/Projects/alpine-phenology/phen_sum_precip_2.csv", sep=",")

nrow(precip.1950_2[precip.1950_2$p.value<0.05,]) #51 -> 45
nrow(precip.1950_2[precip.1950_2$p.value<0.05,])/nrow(precip.1950_2)

#earlier
nrow(precip.1950_2[precip.1950_2$p.value<0.05 & precip.1950_2$Slope<0,]) #2 -> 1
mean(precip.1950_2$Slope[precip.1950_2$p.value<0.05 & precip.1950_2$Slope<0]) 
mean(precip.1950_2$adj.r.squ[precip.1950_2$p.value<0.05 & precip.1950_2$Slope<0])

#later
nrow(precip.1950_2[precip.1950_2$p.value<0.05 & precip.1950_2$Slope>=0,]) #49 -> 44
mean(precip.1950_2$Slope[precip.1950_2$p.value<0.05 & precip.1950_2$Slope>=0])
mean(precip.1950_2$adj.r.squ[precip.1950_2$p.value<0.05 & precip.1950_2$Slope>=0])
```


Figure 4C ??? 
```{r}
ggplot(min1950GDD_2_10, aes(Year, HDD, group=scientificName))+
  geom_point()+
  geom_smooth(method="lm", level=0, lwd = .5, colour = "grey")+
  theme_bw()+
  ylab(expression(GDD))


```

# NPN only precip, high, low
###NPN year     
```{r}
npn.min1950 <- npn.foo
yr.npn_2 <- JulBloomDate.min1950("Year",npn.min1950)

nrow(yr.npn_2) #29 

write.table(yr.npn_2, file = "Q:/Research/Projects/alpine-phenology/phenNPN_yr.csv", sep=",")

nrow(yr.npn_2) 
nrow(yr.npn_2[yr.npn_2$p.value<0.05,])
nrow(yr.npn_2[yr.npn_2$p.value<0.05,])/nrow(yr.npn_2)

#earlier
nrow(yr.npn_2[yr.npn_2$p.value<0.05 & yr.npn_2$Slope<0,]) 
mean(yr.npn_2$Slope[yr.npn_2$p.value<0.05 & yr.npn_2$Slope<0]) 
mean(yr.npn_2$adj.r.squ[yr.npn_2$p.value<0.05 & yr.npn_2$Slope<0])

#later
nrow(yr.npn_2[yr.npn_2$p.value<0.05 & yr.npn_2$Slope>=0,]) #5 -> 0
mean(yr.npn_2$Slope[yr.npn_2$p.value<0.05 & yr.npn_2$Slope>=0])
mean(yr.npn_2$adj.r.squ[yr.npn_2$p.value<0.05 & yr.npn_2$Slope>=0])

```
###NPN high
```{r}
high.npn_2 <- JulBloomDate.min1950("Avg_Hi", npn.min1950)

nrow(high.npn_2) #27 -> 29

write.table(high.npn_2, file = "Q:/Research/Projects/alpine-phenology/phen_NPN_high_2.csv", sep=",")

nrow(high.npn_2[high.npn_2$p.value<0.05,])

#earlier
nrow(high.npn_2[high.npn_2$p.value<0.05 & high.npn_2$Slope<0,]) 
mean(high.npn_2$Slope[high.npn_2$p.value<0.05 & high.npn_2$Slope<0]) 
mean(high.npn_2$adj.r.squ[high.npn_2$p.value<0.05 & high.npn_2$Slope<0])

#later
nrow(high.npn_2[high.npn_2$p.value<0.05 & high.npn_2$Slope>=0,]) 
mean(high.npn_2$Slope[high.npn_2$p.value<0.05 & high.npn_2$Slope>=0])
mean(high.npn_2$adj.r.squ[high.npn_2$p.value<0.05 & high.npn_2$Slope>=0])
```
###NPN low
```{r}
low.npn_2 <- JulBloomDate.min1950("Avg_Lo", npn.min1950)

nrow(low.npn_2) #27 -> 29

write.table(low.npn_2, file = "Q:/Research/Projects/alpine-phenology/phen_NPN_low_2.csv", sep=",")

nrow(low.npn_2[low.npn_2$p.value<0.05,])
nrow(low.npn_2[low.npn_2$p.value<0.05,])/nrow(low.npn_2)

#earlier
nrow(low.npn_2[low.npn_2$p.value<0.05 & low.npn_2$Slope<0,]) 
mean(low.npn_2$Slope[low.npn_2$p.value<0.05 & low.npn_2$Slope<0]) 
mean(low.npn_2$adj.r.squ[low.npn_2$p.value<0.05 & low.npn_2$Slope<0])

#later
nrow(low.npn_2[low.npn_2$p.value<0.05 & low.npn_2$Slope>=0,])
mean(low.npn_2$Slope[low.npn_2$p.value<0.05 & low.npn_2$Slope>=0])
mean(low.npn_2$adj.r.squ[low.npn_2$p.value<0.05 & low.npn_2$Slope>=0])

```
###NPN precip
```{r}
precip.NPN_2 <- JulBloomDate.min1950("Raw_Precip", npn.min1950)

nrow(precip.NPN_2) #27 -> 29

write.table(precip.NPN_2, file = "Q:/Research/Projects/alpine-phenology/phen_NPN_precip_2.csv", sep=",")

nrow(precip.NPN_2[precip.NPN_2$p.value<0.05,]) #1
nrow(precip.NPN_2[precip.NPN_2$p.value<0.05,])/nrow(precip.NPN_2)

#later
nrow(precip.NPN_2[precip.NPN_2$p.value<0.05 & precip.NPN_2$Slope>=0,]) 
mean(precip.NPN_2$Slope[precip.NPN_2$p.value<0.05 & precip.NPN_2$Slope>=0])
mean(precip.NPN_2$adj.r.squ[precip.NPN_2$p.value<0.05 & precip.NPN_2$Slope>=0])

```


Growing Degree Days   
```{r}
results <- lapply(unique(min1950GDD_2_10$ID), function(x){
  if(nrow(min1950GDD_2_10[min1950GDD_2_10$ID==x,])>=10){
  foo1 <- summary(lm(HDD~Year,data=min1950GDD_2_10[min1950GDD_2_10$ID==x,]))
  JYC <- data.frame(cbind(data.frame(t(foo1$coefficients[2,])),
                          foo1$coefficients[1,1],
                          t(foo1$fstatistic),
                          ID = x,
                          unique(min1950GDD_2_10$scientificName[min1950GDD_2_10$ID == x])[1],
                          unique(min1950GDD_2_10$Family.Name[min1950GDD_2_10$ID == x])[1],
                          foo1$adj.r.squared,
                          N = nrow(min1950GDD_2_10[min1950GDD_2_10$scientificName==x,])))
  colnames(JYC) <- c("Slope", "StErr", "t.value", "p.value", "Intercept", 
                   "F Stat", "numDF","denDF",
                   "ID", "Species","Family.Name","adj.r.squ","N")
  JYC
  }
  })


results[[1]]
rst <- do.call(rbind,results)
head(rst)


nrow(rst) #406 -> 377
rst[rst$p.value<0.05,]
nrow(rst[rst$p.value<0.05,])

#earlier
nrow(rst[rst$p.value<0.05 & rst$Slope < 0,])
mean(rst$Slope[rst$p.value<0.05 & rst$Slope < 0]) 
mean(rst$adj.r.squ[rst$p.value<0.05 & rst$Slope < 0]) 

#later
nrow(rst[rst$p.value<0.05 & rst$Slope >= 0,]) 
mean(rst$Slope[rst$p.value<0.05 & rst$Slope >= 0]) 
mean(rst$adj.r.squ[rst$p.value<0.05 & rst$Slope >= 0]) 

write.table(rst,file="Q:/Research/Projects/alpine-phenology/GDD_table5a.csv", sep=",")
```

###NPN Growing Degree Days     
###START HERE#####
```{r}
minNPNGDD_2 <- mergeMinGDD(npn.min1950,tmps)

results.npn <- lapply(unique(minNPNGDD_2$scientificName), function(x){
  if(nrow(minNPNGDD_2[minNPNGDD_2$scientificName==x,])>=10){
  foo1 <- summary(lm(HDD~Year,data=minNPNGDD_2[minNPNGDD_2$scientificName==x,]))
  JYC <- data.frame(cbind(data.frame(t(foo1$coefficients[2,])),
                          foo1$coefficients[1,1],
                          t(foo1$fstatistic),
                          Species = x,
                          foo1$adj.r.squared,
                          N = nrow(minNPNGDD_2[minNPNGDD_2$scientificName==x,])))
  colnames(JYC) <- c("Slope", "StErr", "t.value", "p.value", "Intercept", 
                   "F Stat", "numDF","denDF", "Species", "adj.r.squ","N")
  JYC
  }
  })


results.npn[[1]]
rst.npn <- do.call(rbind,results.npn)
head(rst.npn)


nrow(rst.npn) #22
rst[rst.npn$p.value<0.05,]
nrow(rst.npn[rst.npn$p.value<0.05,])

#earlier
nrow(rst.npn[rst.npn$p.value<0.05 & rst.npn$Slope < 0,])
mean(rst.npn$Slope[rst.npn$p.value<0.05 & rst.npn$Slope < 0]) 
mean(rst.npn$adj.r.squ[rst.npn$p.value<0.05 & rst.npn$Slope < 0]) 

#later
nrow(rst.npn[rst.npn$p.value<0.05 & rst.npn$Slope >= 0,]) 
mean(rst.npn$Slope[rst.npn$p.value<0.05 & rst.npn$Slope >= 0]) 
mean(rst.npn$adj.r.squ[rst.npn$p.value<0.05 & rst.npn$Slope >= 0]) 

rst.npn$genus <- gsub(" .*$", "", rst.npn$Species)
m9 <- merge(rst.npn,asuFams,by="genus")
nrow(m9)  

write.table(rst.npn,file="Q:/Research/Projects/alpine-phenology/GDDNPN_table5a.csv", sep=",")

```

Figure 4C too ????
```{r}
ggplot(minNPNGDD_2, aes(Year, HDD, group=scientificName))+
  geom_point()+
  geom_smooth(method="lm", level=0, lwd = .5, colour = "black")+
  theme_bw()+
  ylab(expression(GDD))
```

#Snow   
Snow report generator <http://wcc.sc.egov.usda.gov/reportGenerator/>   
```{r}
snow <- read.table(text = getURL("https://raw.githubusercontent.com/DenverBotanicGardens/alpine-phenology/master/datasets/Monthly19782016_over3200m.txt"),
                   skip=7, header=FALSE,sep=",", na.strings = " ")


header <- scan(text = getURL("https://raw.githubusercontent.com/DenverBotanicGardens/alpine-phenology/master/datasets/Monthly19782016_over3200m.txt"),skip=5,nlines=1, what=character(),sep=",")
header2 <- scan(text = getURL("https://raw.githubusercontent.com/DenverBotanicGardens/alpine-phenology/master/datasets/Monthly19782016_over3200m.txt"),skip=6,nlines=1,what=character(),sep=",")
names(snow)<-paste(header,header2,sep="_")

head(snow)
```

Average of snow over all stations per month over the years
```{r}
#average over all stations per year
avg.p.yr <- do.call(rbind,lapply(split(snow[,grep("_Snow Depth", names(snow))], 
                                       snow$`Water Year_`), 
                                 function(x) colMeans(x, na.rm = TRUE)))

# get rid of years with no data (every month says 0) - should remove 2016 and Oct and Nov from 1997
avg.p.yr.cc <- avg.p.yr[!is.nan(avg.p.yr[,4]),]
avg.p.yr.cc
# Not to be confused with snowmelt - this is just reformating the data
melt.snow <- melt(avg.p.yr.cc)
head(melt.snow)
melt.snow <- melt.snow[!is.nan(melt.snow$value),]
melt.snow$Var2 <- factor(melt.snow$Var2)
str(melt.snow)


ggplot(melt.snow, aes(as.numeric(Var2),value))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~Var1)
```

Reformat snow data:
```{r}
snow.depth <- grep("_Snow Depth", names(snow))
avg.snow <- aggregate(snow[,c(snow.depth)], by=list(snow$`Water Year_`),
                      function(x) mean(x,na.rm=TRUE))
avg.snow <- avg.snow[!is.na(avg.snow[,5]),]

snow.long <- reshape(avg.snow, direction="long",
                     varying=list(names(avg.snow[,-1])),
                     timevar="Month",
                     times=names(avg.snow[,-1]),
                     idvar=c("Group.1"),
                     v.names="Snow.Depth")


snow.long <- snow.long[!is.nan(snow.long$Snow.Depth),]
snow.long$Month <- factor(snow.long$Month, levels = unique(snow.long$Month))
levels(snow.long$Month)
names(snow.long) <- c("Year","Month","Snow.Depth")
```

```{r}
ggplot(snow.long, aes(Year, Snow.Depth, colour=Month))+
  geom_line()+
  theme_bw()+
  facet_wrap(~Month)+
  geom_smooth(method="lm")
```

## LM of snow depth per month over 1997-2015    
```{r}
summary(aov(Snow.Depth~Year*Month, data=snow.long))

summary(lm(Snow.Depth~Year*Month, data=snow.long))

```

```{r}
yr.snow <- aggregate(snow.long$Snow.Depth, by = list(snow.long$Year),
          function(x) sum(x))
#missing data for 1997 and 2016, take those out
yr.snow <- yr.snow[yr.snow$Group.1 %in% c(1998:2015),]
```

#Correlation of snow depth to precip, high and low   
#Snow depth is averaged over sites and then summed for Oct-Sept water year   
Snow depth is only 1998-2015    
precip (unbloom) is only 1950 - 2011
Overlap is 14 years
```{r}
raw.precip.snow <- merge(unbloom, yr.snow, by.x = "Year", by.y = "Group.1")
colnames(raw.precip.snow) <- c("Year","Precip","HighTemp","LowTemp","SnowDepth")

cor(raw.precip.snow[,-1])
pairs(raw.precip.snow[,-1])
```

Growth traits for tables
```{r}
plantinfo <- read.table("P:/hackathon/alpine-phenology/datasets/USDAplantsdata.txt",
                        header=TRUE, sep = ",")
head(plantinfo)


gtyr <- merge(yr.1950_2, plantinfo, by.x = "Species", by.y = "Scientific.Name")

```
